---
title: "Time Series Project Data Assembly & EDA"
author: "Adhiraj Maheshwari Srivastava"
date: "25/05/2022"
output: html_document
---

This R Markdown consists of this projects EDA. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing Libraries
```{r}
library(forecast)
library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(TSA)
library(Metrics)
library(tidyverse)
library(lubridate)
```

## Reading in the Data
```{r}
wd <- getwd()
wd <- substr(wd, 1, nchar(wd)-5)

filePath <- paste(wd, "/Data/", sep="")
df.Name <- "Final_Data.csv"
holidays.Name <- "holidays_events.csv"

df.group <- read.csv(paste(filePath, df.Name, sep=""))
holidays.df <- read.csv(paste(filePath, holidays.Name, sep=""))
```

# EDA

```{r}
# number of stores: 54
length(unique(df.group$store_nbr))

#total number of sales: 1075377321
sum(df.group$sales)

#total number of transactions: 4683145203
sum(df.group$transaction)

#total number of promotions: 7845496
sum(df.group$onpromotion)

# average sales across all stores: 12848
sales.all=round(mean(df.group$sales), digits = 0)
sales.all

# average number of promotions held across all stores: 94
promotions.all=round(mean(df.group$onpromotion), digits = 0)
promotions.all

# average number of transactions across all stores: 55950
transactions.all=round(mean(df.group$transaction), digits = 0)
transactions.all

# number of days for each stores
print(paste("Number of Rows for Store 1:",nrow(df.group[df.group$store_nbr == 44,]), sep=" "))
print(paste("Number of Rows for Store 2:", nrow(df.group[df.group$store_nbr == 45,]), sep=" "))
print(paste("Number of Rows for Store 3:", nrow(df.group[df.group$store_nbr == 47,]), sep=" "))
print(paste("Number of Rows for Store 4:", nrow(df.group[df.group$store_nbr == 3,]), sep=" "))
print(paste("Number of Rows for Store 5:", nrow(df.group[df.group$store_nbr == 49,]), sep=" "))
```

## Visualizing each of the datasets 

```{r}
plot_data_44= filter(df.group, store_nbr==44)
plot_data_45= filter(df.group, store_nbr==45)
plot_data_47= filter(df.group, store_nbr==47)
plot_data_3= filter(df.group, store_nbr==3)
plot_data_49= filter(df.group, store_nbr==49)
```

```{r}
ggplot(data=plot_data_44, aes(x=as.Date(date), y=sales, group=1)) +
  geom_line()+
  geom_point()

ggplot(data=plot_data_45, aes(x=as.Date(date), y=sales, group=1)) +
  geom_line()+
  geom_point()

ggplot(data=plot_data_47, aes(x=as.Date(date), y=sales, group=1)) +
  geom_line()+
  geom_point()

ggplot(data=plot_data_3, aes(x=as.Date(date), y=sales, group=1)) +
  geom_line()+
  geom_point()

ggplot(data=plot_data_49, aes(x=as.Date(date), y=sales, group=1)) +
  geom_line()+
  geom_point()
```

Plotting the transactions of the top 5 stores across time
```{r}

par(mfrow=c(5,1))

ggplot(plot_data_44, aes(x = date, y = transaction, group=1)) +
  geom_line() + labs(title="Transactions of Store 44")

ggplot(plot_data_45, aes(x = date, y = transaction, group=1)) +
  geom_line() + labs(title="Transactions of Store 45")

ggplot(plot_data_47, aes(x = date, y = transaction, group=1)) +
  geom_line() + labs(title="Transactions of Store 45")

ggplot(plot_data_3, aes(x = date, y = transaction, group=1)) +
  geom_line() + labs(title="Transactions of Store 3")

ggplot(plot_data_49, aes(x = date, y = transaction, group=1)) +
  geom_line() + labs(title="Transactions of Store 49")

```

```{r}
#Setting up the legend for the graph
colors <- c("Store44" = "blue", "Store45" = "red", "Store47" = "green", "Store3" = "purple" , "Store49" = "black")

ggplot() + 
  geom_line(data=plot_data_44, aes(x=date, y=sales, color = "Store44")) + 
  geom_line(data=plot_data_45, aes(x=date, y=sales, color = "Store45")) +
  geom_line(data=plot_data_47, aes(x=date, y=sales, color = "Store47")) +
  geom_line(data=plot_data_3, aes(x=date, y=sales, color = "Store3")) +
  geom_line(data=plot_data_49, aes(x=date, y=sales, color = "Store49")) +
  labs(title="Sales for All Stores", x="Date", y="Sales", color="Legend")+scale_color_manual(values = colors)
```

```{r}
#Setting up the legend for the graph
colors <- c("Store44" = "blue", "Store45" = "red", "Store47" = "green", "Store3" = "purple" , "Store49" = "black")

#Making the graph (Blue for male, and red for female):
ggplot() + 
  geom_line(data=plot_data_44, aes(x=date, y=transactions, color = "Store44")) + 
  geom_line(data=plot_data_45, aes(x=date, y=transactions, color = "Store45")) +
  geom_line(data=plot_data_47, aes(x=date, y=transactions, color = "Store47")) +
  geom_line(data=plot_data_3, aes(x=date, y=transactions, color = "Store3")) +
  geom_line(data=plot_data_49, aes(x=date, y=transactions, color = "Store49")) +
  labs(title="Transactions for All Stores", x="Date", y="Transactions", color="Legend")+scale_color_manual(values = colors)
```

## Calculating basic facts for the top 5 stores
```{r}
# average sales across 5 stores: 31152
sales.5=round(mean(df.group$sales), digits = 0)
sales.5

# average number of promotions held across 5 stores: 112
promotions.5=round(mean(df.group$onpromotion), digits = 0)
promotions.5

# average number of transactions across 5 stores: 117964
transactions.5=round(mean(df.group$transaction), digits = 0)
transactions.5
```

Making a few comparison graphs for avg number sales, avg number, of promotions across stores and the avg number of transactions

```{r}
names=c('All Stores', 'Top 5 Stores in Sales')
sales.comparison=c(sales.all, sales.5)
transaction.comparison=c(transactions.all, transactions.5)
promotion.comparison=c(promotions.all, promotions.5)
```


```{r}
x1=barplot(sales.comparison, names.arg=names, xlab='Stores', ylab='Sales', ylim=c(0, 35000),
        main='Average Number of Sales', col=c('#00203FFF', '#ADEFD1FF'), border = 'black')
text(x1, sales.comparison+1500,labels=as.character(sales.comparison))


x2=barplot(transaction.comparison, names.arg=names, xlab='Stores', ylab='Transactions', ylim=c(0, 130000),
        main='Average Number of Transactions', col=c('#00A4CCFF', '#F95700FF'), border = 'black')
text(x2, transaction.comparison+5000,labels=as.character(transaction.comparison))


x3=barplot(promotion.comparison, names.arg=names, xlab='Stores', ylab='Promotions', ylim=c(0, 150),
        main='Average Number of Promotions', col=c('#949398FF', '#F4DF4EFF'), border = 'black')
text(x3, promotion.comparison+7,labels=as.character(promotion.comparison))
```

## Understanding the different types of holidays

```{r}
holiday_type_counts <- 
  holidays.df %>%                              
  group_by(locale) %>%
  summarise(count = n_distinct(date))
holiday_type_counts 
```

```{r}
x=barplot(holiday_type_counts$count, names.arg=holiday_type_counts$locale, xlab='Type', ylab='Count', ylim=c(0, 250),
        main='Type of Holidays', col=c('#567572FF', '#964F4CFF', '#696667FF'), border = 'black')
text(x, holiday_type_counts$count+10,labels=as.character(holiday_type_counts$count))
```
