---
title: "ARIMA Modeling"
author: "Adhiraj Maheshwari Srivastava"
date: "26/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(forecast)
library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)
library(TSA)
library(Metrics)
library(tidyverse)
library(lubridate)
library(RColorBrewer)
library(xts)
```


Reading in the Data

```{r}
wd <- getwd()
wd <- substr(wd, 1, nchar(wd)-5)

filePath <- paste(wd, "/Data/", sep="")
df.Name <- "Final_Data.csv"

df.group <- read.csv(paste(filePath, df.Name, sep=""))

```

Have created a function below which filters the dataset for the store number that we enter and only selects the date and the sales columns. After this we aggregated the dataset on a weekly level, take the sum of the sales and convert this resultant dataset into a xts object. Next, we split the data into train and test. Test contains the last 10 rows.

We fit an autoarima model on train, use it to predict the next 10 values. This is used to calculate the smape and rmse which are then added into lists.

```{r}
smape_lst=c()
rmse_lst=c()

my_function= function(storenumber, x=df.group){
  
  df=filter(x, store_nbr==as.integer(storenumber))
  df= subset(df, select= c(1,3))
  df_xts=apply.weekly(as.xts(df$sales, order.by=as.Date(df$date)), 'sum')
  train=df_xts["/2017-06-11"]
  test=df_xts["2017-06-18/"]
  my_model=auto.arima(train)
  print(summary(my_model))
  y.pred = forecast(my_model, h=10)$mean
  a=smape(test, y.pred)
  b=rmse(test, y.pred)
  print(paste('smape:', a, 'rmse:', b))
  smape_lst<<-c(smape_lst, a)
  rmse_lst<<-c(rmse_lst, b)
  
}
```

We run the function for all of our stores in the storenumber_lst
```{r}
storenumber_lst=c(44, 45, 47, 3, 49)

for (i in storenumber_lst) {
  my_function(i)
}
```

Combining the results of all the models, the RMSE and sMAPE values in a data frame
```{r}
my_data <- data.frame(storenumber_lst, rmse_lst, smape_lst)
my_data
```

Plotting graphs for the different RMSE and sMAPE values for the models for each of the stores
```{r}
coul = brewer.pal(5, "BuGn") 
x_RMSE=barplot(my_data$rmse_lst, names.arg=my_data$storenumber_lst, xlab='Store', ylab='RMSE',col=coul)
x_RMSE
#text(x, holiday_type_counts$count+10,labels=as.character(holiday_type_counts$count))
x_sMAPE=barplot(my_data$smape_lst, names.arg=my_data$storenumber_lst, xlab='Store', ylab='sMAPE',col=coul)
x_sMAPE
```

Plotting the forecast for Store 49, the store which gives us the least RMSE value
```{r}
plotdf= filter(df.group, store_nbr==49)
plotdf= subset(plotdf, select= c(1,3))
plotdf_xts=apply.weekly(as.xts(plotdf$sales, order.by=as.Date(plotdf$date)), 'sum')
plot.train=plotdf_xts["/2017-06-11"]
plot.test=plotdf_xts["2017-06-18/"]
my_model=auto.arima(plot.train)
y.pred = forecast(my_model, h=10)
autoplot(y.pred, ylab="Store Sales")
```
